---
title: "STAT 385 Midterm Review"
author: "Matthew Hoyle"
date: 2020-03-05T12:00:00-00:00
categories: ["R"]
tags: ["STAT385", "R"]
---

This semester in Stat 385: Statistics Programming Methods I have learned a fair amount of technical skills. Although I entered this class fairly experienced with R, I have learned quite a few R techniques that I was unfimiliar with previously, as well as learning how to use Github Rblogdown.

### R Functions and Techniques

Listed below are a few functions that before entering this course I had either only used sparingly or hadn't used at all. I have thus far have learned a great deal about each of them eitehr through class or personal research.

* cat()
* %in%
* replicate()

\

In addition to these functions, I also have learned a great deal about vectorized coding. Before this class I had never actually intentionally attempted to code vectorized functions. I of course had done so a few times without really understanding what vectorized functions are, since this is the first time I had learned of them.

\

Below is an example of vectorized coding in action, question 1(c) on homework 3. The first chunk of code is the original function, and the second chunk is my vectorized function. As you can see, the vectorized function provides the same result in addition to being shorter both in code as well as run time.

\

Original function:
```{r}
roulette_loop <- function(many_bets) {
  
  # randomly decide whether you lose or win on each bet
  win_lose_random <- sample(x = c(TRUE, FALSE), size = length(many_bets),
                            replace = TRUE)
  
  total_prize <- c()
  
  # for each bet
  for (i in 1:length(many_bets)) {
    bet <- many_bets[i]
    prize <- 0
    
    # if we win, then check how much the prize is
    if (win_lose_random[i]) {
      if (bet == 'low') {
        prize <- 10
      } else if (bet == 'high'){
        prize <- 10
      } else if (bet == 'red') {
        prize <- 20
      } else if (bet == "black") {
        prize <- 20
      } else if (bet == "odd") {
        prize <- 15
      } else if (bet == "even") {
        prize <- 15
      } else if (bet == "first") {
        prize <- 50
      } else if (bet == "second") {
        prize <- 50
      } else if (bet == "third") {
        prize <- 50
      }
    }
  # add prize from this bet to the total prize
  total_prize <- c(total_prize, prize)
  }
  
  total_prize
}
```

```{r}
set.seed(385)
roulette_loop(many_bets = c("red", "black", "low", "high"))
```

```{r}
long_vec <- rep(c("red", "black", "low", "high", "second", "first", "third",
                  "odd", "even"), 10000)
system.time(roulette_loop(many_bets = long_vec))
```
\
Vectorized function:
```{r}
roulette_vec <- function(many_bets){
    
    result <- c()
    win_lose_random <- sample(x = c(TRUE, FALSE), size = length(many_bets),
                              replace = TRUE)
    
    result[many_bets == 'low' | many_bets == 'high' ] <- 10
    result[many_bets == 'red' | many_bets == 'black' ] <- 20
    result[many_bets == 'odd' | many_bets == 'even' ] <- 15
    result[many_bets == 'first' | many_bets == 'second' | many_bets == 'third' ] <- 50
    
    result[win_lose_random == FALSE] <- 0
    
    return(result)
}
```

```{r}
set.seed(385)
roulette_vec(many_bets = c("red", "black", "low", "high"))
```

```{r}
long_vec <- rep(c("red", "black", "low", "high", "second", "first", "third",
                  "odd", "even"), 10000)
system.time(roulette_vec(many_bets = long_vec))
```


